---
title: "Assignment week 4"
author: "Daria Dubovskaia"
output: html_document
---

```{r include=FALSE, warning = FALSE, message = FALSE}
library(dplyr)
library(stringr)
```

### 1. Using the 173 majors listed in fivethirtyeight.com’s College Majors dataset [https://fivethirtyeight.com/features/the-economic-guide-to-picking-a-college-major/], provide code that identifies the majors that contain either "DATA" or "STATISTICS"
Load data
```{r setup, echo=TRUE, results='hide', warning=FALSE, message=FALSE}
majors <- read.csv("https://raw.githubusercontent.com/fivethirtyeight/data/master/college-majors/majors-list.csv")
```
We will use grep function to check if column Major contains word "DATA" or "STATISTICS". 
```{r}
grep_output <- grep('DATA|STATISTICS',majors$Major, value = TRUE)
grep_output
```
Using stringr
```{r}
stringr_output <- majors[str_detect(majors$Major, "STATISTICS|DATA"),]
stringr_output
```

### 2. Write code that transforms the data below:
[1] "bell pepper"  "bilberry"     "blackberry"   "blood orange"
[5] "blueberry"    "cantaloupe"   "chili pepper" "cloudberry"  
[9] "elderberry"   "lime"         "lychee"       "mulberry"    
[13] "olive"        "salal berry"
Into a format like this:
c("bell pepper", "bilberry", "blackberry", "blood orange", "blueberry", "cantaloupe", "chili pepper", "cloudberry", "elderberry", "lime", "lychee", "mulberry", "olive", "salal berry")
I didn't fully get this assignment, should I just create input and make output look like c("bell pepper", etc..)? 
```{r}
input_str <- '[1] "bell pepper"  "bilberry"     "blackberry"   "blood orange"
[5] "blueberry"    "cantaloupe"   "chili pepper" "cloudberry"  
[9] "elderberry"   "lime"         "lychee"       "mulberry"    
[13] "olive"        "salal berry"'
input_str
```
Finding all quoted values:
```{r}
extract_berry <- unlist(str_extract_all(input_str, '"[^"]*"'))
extract_berry 
```
Removing \":
```{r}
extract_berry_update <- str_remove_all(extract_berry, "\"")
extract_berry_update
```
Writing the result in the required formed with c, parentheses
```{r}
final <- paste('c(', paste('"',extract_berry_update,'"',sep = "", collapse = ','), sep = "",')')
str_view(final,'""')
```

### 3.Describe, in words, what these expressions will match:
#### 3.1 (.)\1\1
(.) 1st capturing group, would match any character except for for line terminators  <br>
\1\1 would refer to the same text as previously matched by the first capturing group 2 times <br>
We will need o create a string to use regular expression in R by using \\1 instead of \1.  <br>
For example, 777 in 84572777. In other words, will show 3 characters repeated.
```{r}
stng_1 <- '84572777'
str_view(stng_1, "(.)\\1\\1", match = TRUE)
```

#### 3.2 "(.)(.)\\2\\1"
The regular expression is shown as a string with "" and \\  <br>
(.)(.) 1st and 2nd capturing group, would match any character except for for line terminators  <br>
\\2 would refer to the same text as previously matched by the second capturing group <br>
\\1 would refer to the same text as previously matched by the first capturing group <br>
For example, 4444 in 4564444rt. In other words, will show 4 characters repeated.
```{r}
stng_2 <- '4564444rt'
str_view(stng_2, "(.)(.)\\2\\1", match = TRUE)
```

#### 3.3 (..)\1
(..) would match any 2 characters except for for line terminators <br>
\1 would refer to the same text as previously matched by the first capturing group <br>
For example, 1414 in ssff1414slfs. In other words, will show 2 characters repeated twice.
```{r}
stng_3 <- 'ssff1414slfs'
str_view(stng_3, "(..)\\1", match = TRUE)
```
#### 3.4 "(.).\\1.\\1"
(.) 1st capturing group, would match any character except for for line terminators <br>
. would match any character except for for line terminators/
\\1 would refer to the same text as previously matched by the first capturing group <br>
For example, e7e8e in e7e8e9e7. In other words, will return string if 1st character repeated on the 3rd and 5th positions as well as 2nd and 4th symbols.
```{r}
stng_4 <- 'e7e8e9e7'
str_view(stng_4, "(.).\\1.\\1", match = TRUE)
```

#### 3.5 "(.)(.)(.).*\\3\\2\\1"
(.)(.) 1st,2nd,3rd capturing group, would match any character except for for line terminators  <br>
. would match any character except for for line terminators <br>
* controls how many times a pattern matches between zero and unlimited times, as many times as possible <br>
\\3 would refer to the same text as previously matched by the third capturing group <br>
\\2 would refer to the same text as previously matched by the second capturing group <br>
\\1 would refer to the same text as previously matched by the first capturing group <br>
For example, 1234y4321 in 1234y4321rff. In other words, will return string with the min length of 6 if first three characters are the same as last three in reverse order, there can be any symbols between these two groups.
```{r}
stng_5 <- '1234y4321rff'
str_view(stng_5, "(.)(.)(.).*\\3\\2\\1", match = TRUE)
```

### 4. Construct regular expressions to match words that: 
#### 4.1 Start and end with the same character
To only match a complete string, we should start with ^ and end with $.
```{r}
stng <- c('slow','momnm','slack','hopeh')
str_view(stng, "^(.).*\\1$")
```
#### 4.2 Contain a repeated pair of letters (e.g. "church" contains "ch" repeated twice.)
```{r}
st <- c('slowwew','momnm','scklack','ohopehop','banana')
str_view(st, ".*(.)(.).*\\1\\2.*")
```
#### 4.3 Contain one letter repeated in at least three places (e.g. "eleven" contains three "e"s.)
```{r}
s <- c('slowwew','momnm','scklack','ohopehop','banana')
str_view(s, ".*(.).*\\1.*\\1.*")
```
